Sherlock topic model paper

### 2018-05-18 main.tex
intro framing notes, commented out methods

@@

\subsubsection{Naturalistic extensions of classic memory analyses}

@@

\subsection{Participants and Experimental Design}
How much detail here? or just point to janice's paper?

\subsubsection{Fitting a topic model from text descriptions of Sherlock}
To quantify the flow of information from scene to scene, we used a topic model. A topic model estimates the most likely mixture of topics in a sample of text. First, the video was manually segmented into 1000 scenes, and a collection of descriptive features were manually transcribed.  For each scene, we considered the following features: scene details (i.e. a sentence or two describing what happened in that scene, space (indoor or outdoor), name of all the characters in the scene, name of the character in focus, name of the character speaking, location, camera angle, music presence, and words on the screen. We concatenated the text for all of these features within each segment, creating a 'bag of words' describing each scene. We then transformed the text descriptions into overlapping windows of 50 scene segments. For example, the first text sample comprised the text from the first 50 segments, the second comprised the text from n+1:n+51, and so on. We trained our model using these overlapping text samples using scikit-learn's `CountVectorizer' and `LatentDirichletAllocation' classes.  First, the text was transformed into a vector of word counts (default scikit-learn parameters). This gave a word count vector for each scene in the video.  Then, the word count vectors were used to fit a topic model (number of topics=100, method=batch).

\subsubsection{Projecting the video and verbal recall into a common space}
We used the model described above to transform the video text descriptions and verbal recall transcripts into a common space.  We created the video model by transforming exactly the same text (overlapping scene descriptions) with the topic model, resulting in a scene (1000) by topic (100) matrix that summarized the topics in each scene (Fig. \ref{fig:model}a). The scene descriptions often spanned multiple timepoints (i.e. TRs). To account for this, we expanded the video model by copying the rows of the model for as many timepoints that the scene description spanned. After this expansion, the shape of the model was the length of the duration of the video (1976 TRs).

To create the recall models, for each participant we tokenized the recall transcript into a list of sentences and then mapped the list to overlapping windows of 10 sentences.  We transformed the list of overlapping recall sentences using the model that was trained on the video text. The result of this was a matrix (\# of sentences by 100 topics) that represented the most likely mixture of topics for a given chunk of sentences. We resampled the recall model to be the same length as the video model (1976 samples). The result was 17 participant-specific recall models (1976 timepoints by 100 topics). Intuitively, if the recall transcript for a given participant correctly described each video scene in order, the video and recall models should match closely.  However, if the scenes were recalled out of order, omitted, or described using language that was very different from the video scene descriptions, the video and recall models would look very different. Figure \ref{fig:model}b shows the group average recall model. Qualitatively, the two matrices appear to be similar. Our next set of analyses sought to quantify this relationship.

To quantify the similarity between the movie and recall models, for each participant we correlated every moment of the video model with every moment of the recall models, resulting in a timepoints (1976) by timepoints (1976) correlation matrix for each participant.  The individual matrices for each participant can be seen in Supp Fig. X and the group average is displayed in \ref{fig:model}c.

\subsubsection{Estimating the number of recall events for each participant}
Our metric for choosing k was to select the value that maximized the ratio of the average `within-event' correlation (i.e. blocks of time the model identified as an event) and the average `across-event' correlation.

\subsubsection{Characterizing memory performance}
We used a few different metrics to assess memory that were all designed to capture the relationship between the video and recall models.  First, to get a general sense for the match between the video and recall matrices, we vectorized the the matrices and correlated them.  This resulted in a single number representing how well a given participant recalled the video. To get a temporally dynamic measure of memory, for each participant we computed the pairwise correlation between matching video/recall timepoints.  This allowed us to assess memory for individual moments of the video. Finally to get a more nuanced representation of memory for the video, we computed a video/recall model correlation matrix for each participant. These timepoint (1976) by timepoint (1976) matrices represent the correlation between every moment of the video model and every moment of the recall model. To provide some intuition, if a participant recalled every scene at the same rate and in exactly the same order as the video, we would expect high correlation values along the diagonal of this matrix. However, if a participant recalls the scenes out of order, or at different rates for different scenes, this will result in high values off the diagonal.

>>>>

- using list-learning / trial based experiments, we've learned a great deal about the human memory system.
- however, its undeniable that many lab-based experiments are severely contrived as compared to real life experiences
- for example, real life experiences are typically highly autocorrelated in space and time, a feature that is not typically present in traditional laboratory studies.
- nonetheless, we want to leverage the tools and knowledge gained over the past few decades
- here, we present a methodological and theoretical advance in our understanding of memory for naturalistic experiences.
- our approach is model the semantic information contained in a stimulus and project memory responses into this stimulus space, allowing for fine-grained analysis of the mnemonic reprsentations.

@@

\subsubsection{Naturalistic extensions of classic memory analyses}
s
Traditionally, a major limitation of 'naturalistic' memory experiments has been quantifying memory dynamics in a principled way.
In the next section, we extend a set of classic memory analyses that were originally designed to measure memory in 'free recall' list-learning experiments to naturalistic settings. Just like recall

@@

% \subsection{Participants and Experimental Design}
% How much detail here? or just point to janice's paper?
%
% \subsubsection{Fitting a topic model from text descriptions of Sherlock}
% To quantify the flow of information from scene to scene, we used a topic model. A topic model estimates the most likely mixture of topics in a sample of text. First, the video was manually segmented into 1000 scenes, and a collection of descriptive features were manually transcribed.  For each scene, we considered the following features: scene details (i.e. a sentence or two describing what happened in that scene, space (indoor or outdoor), name of all the characters in the scene, name of the character in focus, name of the character speaking, location, camera angle, music presence, and words on the screen. We concatenated the text for all of these features within each segment, creating a 'bag of words' describing each scene. We then transformed the text descriptions into overlapping windows of 50 scene segments. For example, the first text sample comprised the text from the first 50 segments, the second comprised the text from n+1:n+51, and so on. We trained our model using these overlapping text samples using scikit-learn's `CountVectorizer' and `LatentDirichletAllocation' classes.  First, the text was transformed into a vector of word counts (default scikit-learn parameters). This gave a word count vector for each scene in the video.  Then, the word count vectors were used to fit a topic model (number of topics=100, method=batch).
%
% \subsubsection{Projecting the video and verbal recall into a common space}
% We used the model described above to transform the video text descriptions and verbal recall transcripts into a common space.  We created the video model by transforming exactly the same text (overlapping scene descriptions) with the topic model, resulting in a scene (1000) by topic (100) matrix that summarized the topics in each scene (Fig. \ref{fig:model}a). The scene descriptions often spanned multiple timepoints (i.e. TRs). To account for this, we expanded the video model by copying the rows of the model for as many timepoints that the scene description spanned. After this expansion, the shape of the model was the length of the duration of the video (1976 TRs).
%
% To create the recall models, for each participant we tokenized the recall transcript into a list of sentences and then mapped the list to overlapping windows of 10 sentences.  We transformed the list of overlapping recall sentences using the model that was trained on the video text. The result of this was a matrix (\# of sentences by 100 topics) that represented the most likely mixture of topics for a given chunk of sentences. We resampled the recall model to be the same length as the video model (1976 samples). The result was 17 participant-specific recall models (1976 timepoints by 100 topics). Intuitively, if the recall transcript for a given participant correctly described each video scene in order, the video and recall models should match closely.  However, if the scenes were recalled out of order, omitted, or described using language that was very different from the video scene descriptions, the video and recall models would look very different. Figure \ref{fig:model}b shows the group average recall model. Qualitatively, the two matrices appear to be similar. Our next set of analyses sought to quantify this relationship.
%
% To quantify the similarity between the movie and recall models, for each participant we correlated every moment of the video model with every moment of the recall models, resulting in a timepoints (1976) by timepoints (1976) correlation matrix for each participant.  The individual matrices for each participant can be seen in Supp Fig. X and the group average is displayed in \ref{fig:model}c.
%
% \subsubsection{Estimating the number of recall events for each participant}
% Our metric for choosing k was to select the value that maximized the ratio of the average `within-event' correlation (i.e. blocks of time the model identified as an event) and the average `across-event' correlation.
%
% \subsubsection{Characterizing memory performance}
% We used a few different metrics to assess memory that were all designed to capture the relationship between the video and recall models.  First, to get a general sense for the match between the video and recall matrices, we vectorized the the matrices and correlated them.  This resulted in a single number representing how well a given participant recalled the video. To get a temporally dynamic measure of memory, for each participant we computed the pairwise correlation between matching video/recall timepoints.  This allowed us to assess memory for individual moments of the video. Finally to get a more nuanced representation of memory for the video, we computed a video/recall model correlation matrix for each participant. These timepoint (1976) by timepoint (1976) matrices represent the correlation between every moment of the video model and every moment of the recall model. To provide some intuition, if a participant recalled every scene at the same rate and in exactly the same order as the video, we would expect high correlation values along the diagonal of this matrix. However, if a participant recalls the scenes out of order, or at different rates for different scenes, this will result in high values off the diagonal.
